\ifdefined\MAINTEXT \else Supplemental Material: \\ \fi Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2017-10-01 almo-langevin.tex
No summary, no abstract yet

@@

\title{
Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
%%for weakly-coupled atoms
%% for non-covalently bonded atoms
}

@@

@@

The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.

@@

@@

Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
\emph{ab initio} molecular dynamics (AIMD)~\cite{b:aimd} has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids. %[of gas-phase molecules and condensed-phase systems.] 
Unfortunately, the computational cost of the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms, which severely limits the \emph{length scales} accessible by AIMD. %and hinders it application to large heterogeneous systems (interfaces, nuclei, defects). 
To address this issue and calculate forces in large systems efficiently, substantial efforts have been directed to the development of linear scaling (LS) DFT methods.

% Re-write the following section and include any recent advances
In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian, known as canonical KS orbitals, must be replaced with an alternative set of \emph{local} electronic descriptors
~\footnote{Canonical KS orbitals are delocalized over all atoms in the system. 
This delocalization is the main reason for the steep growth of the cost of the conventional KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom that describe delocalized orbitals grows quadratically.}. 
Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM). 
However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis. 
Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.

To obviate the DM optimization, another set of natural local electronic descriptors can be constructed by enforcing strict locality on the KS orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are very promising since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the development of orbital-based LS methods has been hindered by the inherently difficult optimization of localized orbitals reported by multiple research groups~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.

%ZZZ-important addition: we did not include a large family of the so-called fragmentation methods 
%included only those methods that attempt to construct  set 
%It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}. These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost. We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix. This approach provides the most rigorous description of the electronic structure.

Thus, despite remarkable recent progress the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets. 
On the length and time scales required for accurate routine AIMD simulations, LS DFT cannot compete with the straightforward low-cost cubically-scaling KS DFT.

In this work, we introduce a new orbital-based AIMD method that is not only linear scaling but also has a low computational overhead. 
Currently, the method is restricted to large ensembles of atoms or small molecules that do not form strong covalent bonds with each other. 
Using liquid water as an example, we demonstrate that the new AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center and a predefined localization radius $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is a set of all local Gaussian atomic orbitals of one molecule. 
However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Galerkin}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed variationally in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization that have rendered previous MO-based LS methods impractical. 
In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second step, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first step. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

@@

%For any finite value of $R_{c}$, forces are undefined at manifold of due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations. 
%Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated. 
%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{ZZZ-any-references}. This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.

@@

In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense non-SCF component of the forces. 
The difference between these approximate ALMO forces and the \emph{reference} forces that would be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
%
\begin{align}
\label{eq:deltaf}
f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
\end{align}
%
where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$. $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization. 
$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and decreasing $\epsilon_{\text{SCF}}$. %$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
%ZZZ For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.

@@

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{doi:10.1002/wcms.1176}. Inspired by this method, ALMO AIMD is chosen by be governed by the Langevin equations of motion, which can be written using the unknown reference forces as:
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is stochastic force represented by a white Gaussian noise with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
%
This relation means that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms will maintain a constant average temperature $T$ and generate the correct Maxwell-Boltzmann distribution of velocities~\cite{a:Kubo-1986,b:tuckerman-stat}. In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled. Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to more random noise in trajectories, faster equilibration but more distorted results for calculated dynamical properties.

@@

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by a stochastic white Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha} = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This assumption, shown below to be well justified (Figure~\ref{fig:randomforce}), allows us to write the Langevin equations using the approximate ALMO forces:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. This new Langevin equation means that the value of $\Delta$ must be known to re-balance the damping and stochastic terms and generate the correct Maxwell-Boltzmann distribution in simulations.

Note that $\Delta$ can be either positive or negative depending on the way errors are introduced in the system of interest. 
$\Delta$ was found to be positive for all systems considered here, increasing the amplitude of the effective random force on the system.

%\section{Optimizing the values of $\gamma$ and $\Delta$}

In theory, both $\gamma$ and $\Delta$ should be system-size independent, but in practice only $\gamma$ is.
We found that $\Delta$ is dependent on system size due to finite size effects, which decay for increasing system size.
Fortunately, $\gamma$ and $\Delta$ can be separately optimized for a given system of interest, assuming a method of accurate force calculation can be applied to the system for a short trajectory.

An optimized $\gamma$ is one that is: large enough to adequately absorb force discontinuities introduced by finite $R_{c}$; large enough to correctly thermostat the system of interest on a reasonable time scale; and small enough not to significantly affect dynamical properties on the time scale of interest. 
The relative weight of the three is determined by the specifics of the system and what is most important to the researcher. 
To optimize $\gamma$ given the relative importance of the above properties, we suggest running several short trajectories on the system of interest (or a scaled-down version of it) using a tightly converged potential energy surface calculation and varying the value of $\gamma$. 
We found that a value of $\gamma$ between $10^{-5}$ and $10^{-3}\ \mathrm{fs^{-1}}$ was optimal for the purposes of the research presented here.

To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.

$R^{-\Delta}_{i\alpha}$ can therefore be optimized simply by trying various values of $\Delta$ (with \mbox{$\gamma = 0$}) until the simulation has long-term stable total energy and satisfies the equipartition theorem. 
%
\begin{align}
\label{eq:eqipartition}
\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
\end{align}
%

Another approach based on Eq. (\ref{eq:assumption}) can be used to find $\Delta$ within an order of magnitude if one can afford computing tightly converged SCF forces for a short MD trajectory. 
First, run the simulation with $\Delta = 0$ and converged SCF forces. 
Second, compute approximate forces for snapshots along the trajectory and calculate the random term $R^{\Delta}_{i\alpha} (t)$ using Eq.\ \ref{eq:assumption}. 
Third, calculate the autocovariance function (ACF) of the noise $\text{ACF}(t) = \langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle$, which should be sharply peaked around zero lag time if our assumption is correct. 
Finally, integrate Eq.(\ref{eq:stochastic3}) to obtain the following expression for $\Delta$.
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle R^{\Delta}_{i\alpha} (0)  R^{\Delta}_{i\alpha} (t) \rangle dt
\end{align}
%
In other words, we use the integral of the ACF to estimate the strength of the random force. 
In practice, we found this method for estimating $\Delta$ was not particularly accurate due to rapidly oscillating correlations between sequential values of $R^{\Delta}_{i\alpha} (t)$. 
The ACF was not possible to accurately integrate without running a reference trajectory with extremely short time-steps (see figure~\ref{fig:randomforce}).
For example, using Eq.\ \ref{eq:delta} we calculated $\Delta \approx \SI{2e-5}{\per\fs}$, but the value was optimized heuristically to $\SI{6e-5}{\per\fs}$.
Still, this method provided a decent order of magnitude starting point for further investigation.
Note that for our system, the ACF decays rapidly enough that the forces can be considered random and uncorrelated to a good approximation for sufficiently long time scales ($\gg\SI{200}{\fs}$).

\begin{figure}
\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_with_ACF.eps}
\caption{\label{fig:randomforce} Comparison of random force introduced from $R_{c}$ and $\epsilon_{\text{SCF}}$. 
The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line. 
(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged ALMO SCF as a reference.
(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, using fully converged OT SCF as a reference.
(c) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., using fully converged OT SCF as a reference.
(d) Normalized ACF of the instantaneous error in panel (a) $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it} $.
Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
}
\end{figure}

@@

The methods presented here are built into CP2K, an open source MD simulation package.~\cite{www:cp2k} 
The computation scheme used is mostly identical to that described in the implementation section of ref.\ \citenum{a:almo-ls}. 
For the dynamical properties calculations using ALMO DFT methods, a periodic cell containing 125 water molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{293}{\K}$ and a constant density of $\SI{1.00597}{\g\per\cm^{3}}$. 
The system was thermostated using the Langevin ensemble (eq.\ \ref{eq:langevin2}) where $\gamma$ was set to $\SI{e-3}{\per\fs}$, and $\Delta$ was optimized to the order of $10^{-5}\, \mathrm{fs^{-1}}$. 
Note that for timing benchmarks, $\Delta$ was not optimized. 
This does not have a perceptible effect on calculation time.

Simulations for timing benchmarks were run for ten time steps, which is sufficient for convergence of wavefunctions from an initial guess.
The TZV2P basis set was used to represent MOs, and a cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The exchange-correlation energy was approximated using the PBE functional, which expands upon the local density approximation by considering derivatives of the electron density at each point.~\cite{a:PBEfunctional}

The reference electronic structure was calculated using the orbital transformation (OT) method~\cite{a:ot,a:ot2} at $\SI{302}{\K}$ and the Langevin thermostat with $\gamma = \SI{e-3}{\per\fs}$. 
All other system parameters were set identical to the ALMO system.

IR spectra, radial distribution functions, diffusivity plots, and diffusion constants were calculated using the open source trajectory analyzer program TRAVIS.~\cite{a:travis-main,a:travis-ir1,a:travis-ir2}

>>>>

% define variable that compiles the main text (as opposed to SI text)
\newcommand*{\MAINTEXT}{}

@@

\ifdefined\MAINTEXT
\else
	\clearpage
	\widetext
	\setcounter{figure}{0}
	\setcounter{page}{1}
	\renewcommand{\thefigure}{S\arabic{figure}}
\fi

\title{
\ifdefined\MAINTEXT
\else
Supplemental Material: \\
\fi
Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
%%for weakly-coupled atoms
%% for non-covalently bonded atoms
}

@@

\ifdefined\MAINTEXT

@@

%The approach utilizes a systematically inaccurate potential energy surface calculation method that grows linearly in computational complexity with the size of the simulated system. 
%Systematic inaccuracies in the calculation of force using this method are balanced by a weighted version of the Langevin equation of motion. 
%The stochastic ``white noise'' of the Langevin equation acts to absorb force discontinuities while also thermostating the system to maintain the desired thermodynamic temperature. 
%In addition, we introduce a simple heuristic approach to find optimized Langevin equation parameters to use for balancing system-specific imperfect forces and maintaining correct dynamical behavior.

@@

\fi

@@

Since the unification of molecular dynamics and density functional theory (DFT)~\cite{a:thecpmd},
\emph{ab initio} molecular dynamics (AIMD) has become an important tool for the study of weakly bonded ensembles of molecules such as gas-phase molecular clusters, liquids, solutions, and molecular solids. %[of gas-phase molecules and condensed-phase systems.] 
Unfortunately, the computational cost of the conventional Kohn-Sham (KS) DFT grows cubically with the number of atoms, which severely limits the \emph{length scales} accessible by AIMD. %and hinders it application to large heterogeneous systems (interfaces, nuclei, defects). 
To address this issue, substantial efforts have been directed to the development of linear scaling (LS) DFT methods.

% Re-write the following section and include any recent advances
In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors
~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}. 
Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM). 
However, in typical DFT calculations that require diffuse basis functions, the DM approach is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003, RZK-MAO} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis. 
Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.

To obviate the DM optimization, another set of electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the development of orbital-based LS methods has been all but abandoned because of the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.

%RZK-important addition: we did not include a large family of the so-called fragmentation methods 
%included only those methods that attempt to construct  set 
%It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}. These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost. We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix. This approach provides the most rigorous description of the electronic structure.

Thus, the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets. 
On the length and time scales accessible in typical AIMD simulations, LS DFT cannot compete with the straightforward low-cost cubically-scaling KS DFT.

In this work, we developed a new orbital-based AIMD method that is not only linear scaling but also has a low computational overhead. 
Currently, the method is restricted to systems of small molecules that do not form strong covalent bonds with each other. 
The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center and a predefined localization radius $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is a set of all local Gaussian atomic orbitals of one molecule. 
However the approach can be easily generalized to other local and nonlocal basis sets~\cite{a:ls-galli-parrinello-1992, Galerkin}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second step, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first step. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

@@

%For any finite value of $R_{c}$, forces are undefined at manifold of due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations. 
%Setting $R_{c}$ to larger values can reduce the magnitude of the discontinuity, but increases the computational cost as longer distance (but physically less important) interactions must be calculated. 
%As long as the discontinuity in force is small compared with the random numerical noise in the calculation, it can be safely ignored~\cite{RZK-any-references}. This means that $R_{c}$ has to be kept relatively large to maintain stable and accurate dynamics.

@@

In our approach, the forces on atoms are calculated \emph{approximately} after the two-step ALMO SCF using a straightforward procedure that computes only the Hellmann-Feynman and Pulay components and neglects the computationally intense non-SCF component of the forces. 
The difference between these approximate ALMO forces and the \emph{reference} forces that would be obtained from perfectly converged fully-delocalized KS orbitals is $\delta f_{i\alpha}(t)$:
%
\begin{align}
\label{eq:deltaf}
f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
\end{align}
%
where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time $t$. $\delta f_{i\alpha} (t)$ comprises all neglected terms that originate from a finite localization radius $R_c$ and incomplete SCF optimization. 
$\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ and decreasing $\epsilon_{\text{SCF}}$. %$\epsilon_{\text{SCF}}$ is chosen such that the electronic ground state is considered sufficiently close to the Born-Oppenheimer state for values less than $\epsilon_{\text{SCF}}$.
%RZK For $\epsilon_{\text{SCF}} \leqslant 10^{-7}$~a.u. ALMO SCF is considered converged.

@@

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term follows the methodology that have been introduced into AIMD by Krajewski \emph{et al.}~\cite{PRB-73-041105R}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{doi:10.1002/wcms.1176}. 
Inspired by this method, ALMO AIMD is chosen by be governed by the Langevin equations of motion, which can be written using the unknown reference forces as:
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is stochastic force represented by a white Gaussian noise with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
%
This relation means that, for any value of $\gamma$, the perfectly balanced damping and stochastic terms will maintain a constant average temperature $T$ and generate the correct Maxwell-Boltzmann distribution of velocities~\cite{a:Kubo-1986,b:tuckerman-stat}. 
In the limit $\gamma \rightarrow 0$, the Newton equations of motion are recovered and the microcanonical ensemble is sampled. 
Finite values of $\gamma$ correspond to imposing canonical thermostat on the system, with larger $\gamma$ leading to more random noise in trajectories, faster equilibration but more distorted results for calculated dynamical properties.

@@

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This assumption, shown below to be well justified, allows us to rewrite the Langevin equations using the ALMO forces:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
This term compensates for the imperfection of ALMO forces and is required to re-balance the damping and stochastic components in the Langevin equation and generate the correct Maxwell-Boltzmann distribution in ALMO AIMD.

%%Note that $\Delta$ can be either positive or negative depending on the system of interest. 
%%$\Delta$ was found to be positive for all simulations considered here. %, increasing the amplitude of the effective random force on the system.

%\section{Optimizing $\Delta$}

%To correctly thermostat the system of interest, $\Delta$ needs to be reduced to zero.
%This is done by finding the value of $\Delta$ and adding a third random force term $R^{-\Delta}_{i\alpha}$ to equation \ref{eq:langevin2} which cancels out the extra forces introduced to the system by $R^{\Delta}_{i\alpha}$.
%This means that a trajectory with inaccurate forces and optimized $R^{-\Delta}_{i\alpha}$ should have no long term drift in the total energy of the system. 
%Short term energy fluctuations should be expected, and can later be ``smoothed out'' by setting $\gamma$ as needed.

(RZK the integral contains dummy index $i$ in the r.h.s.) In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3})
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty}\langle \delta f_{i\alpha} (0)  \delta f_{i\alpha} (t) \rangle dt
\end{align}
%
if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory (see Supplemental Material). 
In practice, we found (see results below) that this approach is not particularly accurate due to the noisy autocovariance functions (ACF) from short trajectories and perhaps because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold. 
Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$. 
This or, in fact, any starting value can be  rapidly fine-tuned in a series of trial-and-error short ALMO AIMD runs until the equipartition theorem is satisfied: 
%
\begin{align}
\label{eq:eqipartition}
\langle \dfrac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle &= \dfrac{3}{2} k_{B} T
\end{align}
%
or, in other words, until the simulation stabilizes with the correct average temperature.

@@

ALMO AIMD presented here was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly in approximate ALMO AIMD to reproduce static and dynamical properties of liquid water. 
%To calculate properties of water at ambient conditions, 
A periodic cell containing 125 molecules was simulated for $\SI{10}{\ps}$ at a temperature of $\SI{298}{\K}$ and a constant density of $\SI{1.006}{\g\cdot\cm^{-3}}$. 
% RZK: we do not need to specify details on gamma optimization since they are well-known.
RZK algorithm~\cite{RZK} was used to integrate the Langevin equation of motion. We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system on the time scale of the simulation and small enough not to significantly affect dynamical properties of liquid water.
%
In the dual-representation Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The XC energy was approximated using the PBE functional~\cite{a:PBEfunctional}. 
The Brillouin zone was sampled at the $\Gamma$-point and separable norm-conserving pseudopotentials were used to describe the interactions between the valence electrons and the ionic cores~\cite{a:hgh}.
